---
title: "AHS-2 Medicare Linkage"
output: github_document
---

```{r setup, include=FALSE}
# Required libraries
pacs <- c("tidyverse", "readxl", "lubridate", "tableone")
sapply(pacs, require, character.only = TRUE)
```

## Datasets 

* Data received from CMS on October 31, 2022
* Contains 4 self-decrypting EXE files for each year of 2008-2020
  * Master Beneficiary Summary File (MBSFABCD)
  * Chronic Conditions (MBSFCC)
  * Cost and Utilization (MBSFCU)
  * MEDPAR (MEDP)
* For 2017 and later, we have:
  * An additional file for a new set of 30 chronic conditions (MBSFCHRONIC)
* Also includes a crosswalk file between SSN and BeneID

## Crosswalk
### SSN matches

```{r read_xwalk, echo = FALSE, message = FALSE}
# Read crosswalk file: n = 70.968
crosswalk <- read_fwf("./Data/12172/2022/ssn_bene_xwalk_res000058038_req012172_2022.dat",
                      fwf_widths(c(9, 15, 1, 1, 1), c("ORIG_SSN", "BENE_ID", "SSN_MATCH", "SEX_MATCH", "DOB_MATCH")))
```

* The crosswalk file contains `r nrow(crosswalk)` SSNs (``ORIG_SSN``).
* There are 3 matching indicators:
  * ``SSN_MATCH`` for SSN-to-BeneID match
  * and whenever there is a match, we also have ``SEX_MATCH``, and ``DOB_MATCH``
  
* Among 70,968 subjects, there were 52,704 matches (74%):
  * There were 115 invalid SSNs (``SSN_MATCH = 9``)

```{r xwalk_ssn_match, echo = FALSE}
crosswalk %>% 
  group_by(SSN_MATCH) %>% 
  tally() %>% 
  mutate(pct = n / sum(n) * 100)
```

* Among 52,704 matches, there were 824 subjects (1.6%) whose gender mismatched with AHS-2 data
  * How should we deal with these gender mismatches? -- to be removed.

```{r xwalk_sex_match, echo = FALSE}
# Gender mismatch -- 824 subjects (1.6%)
crosswalk %>% 
  filter(SSN_MATCH == 1) %>% 
  group_by(SEX_MATCH) %>% 
  tally() %>% 
  mutate(pct = n / sum(n) * 100)
```

* Among 52,704 matches, there were 3331 subjects (6.3%) whose DOB mismatched with AHS-2 data
  * How should we deal with these DOB mismatches? -- to be removed.

```{r xwalk_dob_match, echo = FALSE}
# DOB mismatch -- 3331 subjects (6.3%)
crosswalk %>% 
  filter(SSN_MATCH == 1) %>% 
  group_by(DOB_MATCH) %>% 
  tally() %>% 
  mutate(pct = n / sum(n) * 100)

```

### Anomalies

* There were two BeneIDs that appeared twice in the crosswalk file. 
  * Each of these BeneIDs have two different SSNs (below shows only the last 4 digits of SSN)

```{r xwalk_beneID_duplicates, echo = FALSE}
# Extract matched BENE_IDs
all_matched_bene_ids <- crosswalk %>% 
  filter(SSN_MATCH == 1) 

# BENE_IDs of gender or DOB mismatch (to be removed)
mismatches <- crosswalk %>% 
  filter(SSN_MATCH == 1 & (SEX_MATCH == 0 | DOB_MATCH == 0)) %>%
  select(BENE_ID)

dup_BENE_IDs <- all_matched_bene_ids %>% 
  group_by(BENE_ID) %>% 
  summarize(n = sum(n())) %>% 
  filter(n > 1)

all_matched_bene_ids %>% 
  filter(BENE_ID %in% dup_BENE_IDs$BENE_ID) %>% 
  mutate(SSN_hidden = substr(ORIG_SSN, 6, 10)) %>% 
  select(SSN_hidden, BENE_ID, SSN_MATCH, SEX_MATCH, DOB_MATCH)
```

* There were 106 SSNs that appeared twice in the crosswalk file -- see the firse several examples below
  * Each of these SSNs have two different BeneIDs  (below shows only the last 4 digits of SSN)

```{r xwalk_ssn_duplicates, echo = FALSE}
dup_SSNs <- all_matched_bene_ids %>% 
  group_by(ORIG_SSN) %>% 
  summarize(n = sum(n())) %>% 
  filter(n > 1)

all_matched_bene_ids %>% 
  filter(ORIG_SSN %in% dup_SSNs$ORIG_SSN) %>% 
  mutate(SSN_hidden = substr(ORIG_SSN, 6, 10)) %>% 
  select(SSN_hidden, BENE_ID, SSN_MATCH, SEX_MATCH, DOB_MATCH)
```

```{r exclude_beneids, echo = FALSE, message = FALSE}
# BENE_IDs that need to be removed:
# Gender/DOB mismatch
# SSN/BENE_ID duplicates
exclude_BENE_IDs <- dup_BENE_IDs %>% 
  select(BENE_ID) %>% 
  union(all_matched_bene_ids %>% 
          filter(ORIG_SSN %in% dup_SSNs$ORIG_SSN) %>% 
          select(BENE_ID)
  ) %>% 
  union(mismatches)
```

* These duplicates (n = 212) are to be excluded for analysis below. Similarly, any BeneIDs with gender or DOB mismatch are to be removed (n = 3368).
  * In total, `r length(unique(exclude_BENE_IDs$BENE_ID))` BeneIDs are to be removed.

## Master Beneficiary Summary (MBSF) File

```{r read_msbf, echo = FALSE, message = FALSE}
# Read MBSF Summary data on each year
# Data specification of MSBF files
fts_msbf <- read_excel("./Data/mbsf_format.xlsx")

# Create file names
year <- 2008:2020
fname <- paste0("./Data/12172/", year, "/mbsf_abcd_summary_res000058038_req012172_", year, ".dat")

# Read all MSBF files of 13 years
all_msbf <- fname %>% 
  lapply(\(x) read_fwf(x, fwf_widths(fts_msbf$length, fts_msbf$long_name))) %>% 
  lapply(\(x) anti_join(x, exclude_BENE_IDs)) %>% 
  setNames(year)
```

* All MBSF files of 2008-2020 were imported into R. After excluding duplicated cases described above, the number of observations each year is:

```{r msbf_nobs, echo = FALSE}
all_msbf %>% sapply(nrow)
```

```{r msbf_long, echo = FALSE, message = FALSE}
# Long format over years
all_msbf_long <- all_msbf %>% 
  do.call(rbind, .) %>% 
  arrange(BENE_ID, BENE_ENROLLMT_REF_YR)

# Unique beneficiaries across years: n = 44,585
all_msbf_bene_ids <- all_msbf_long %>% 
  select(BENE_ID) %>% 
  distinct()

# 4546 matched beneficiaries were never appeared in MSBF files
bene_ids_no_show <- all_matched_bene_ids %>% 
  anti_join(exclude_BENE_IDs) %>% 
  anti_join(all_msbf_bene_ids)
```

* After combining all years of data, we have `r nrow(all_msbf_bene_ids)` unique BeneIDs in these 13 years.
  * This indicates that there are `r nrow(bene_ids_no_show)` matched beneficiaries that do not appear in the MSBF files.

### Age, gender and deaths

* See the distribution of age at last seen. Notice the beneficiaries of aged <60 years.

```{r msbf_age_dist, echo = FALSE}
# Age at the last year of appearance
all_msbf_long %>% 
  group_by(BENE_ID) %>% 
  slice(n()) %>% 
  mutate(AgeCat = cut(AGE_AT_END_REF_YR, breaks = c(0, 3:12 * 10), right = FALSE)) %>% 
  group_by(AgeCat) %>% 
  tally() %>% 
  mutate(pct = n / sum(n) * 100)
```

* Current reason for Medicare entitlement (at last seen) indicates that 3.5% of all beneficiaries are because of disability and/or end-stage renal disease:

```{r msbf_entitle, echo = FALSE}
# Current Reason for Entitlement Code
all_msbf_long %>% 
  group_by(BENE_ID) %>% 
  slice(n()) %>% 
  mutate(ENTLMT_RSN_CURR = factor(ENTLMT_RSN_CURR, levels = 0:3, labels = c("OASI", "DIB", "ESRD", "DIB & ESRD"))) %>% 
  group_by(ENTLMT_RSN_CURR) %>% 
  tally() %>% 
  mutate(pct = n / sum(n) * 100)
```

* See the distribution of sex according to the year last seen:

```{r msbf_sex_dist, echo = FALSE}
# Sex at the last year of appearance
all_msbf_long %>% 
  group_by(BENE_ID) %>% 
  slice(n()) %>% 
  group_by(SEX_IDENT_CD) %>% 
  tally() %>% 
  mutate(pct = n / sum(n) * 100)
```

* How many of them died during 2008-2020? There were 13,218 (29.6%) deaths during the 13 years:

```{r msbf_deaths, echo = FALSE}
all_msbf_long %>%
  group_by(BENE_ID) %>% 
  slice(n()) %>% 
  mutate(Dead = ifelse(is.na(BENE_DEATH_DT), 0, 1),
         Dead = factor(Dead, levels = 0:1, labels = c("Alive", "Dead"))) %>% 
  group_by(Dead) %>% 
  tally() %>% 
  mutate(pct = n / sum(n) * 100)
```

* See the distribution of deaths by years:

```{r msbf_deaths_dist, echo = FALSE}
all_msbf_long %>%
  group_by(BENE_ID) %>% 
  slice(n()) %>% 
  mutate(year_died = substr(BENE_DEATH_DT, 1, 4)) %>% 
  group_by(year_died) %>% 
  tally() %>% 
  mutate(pct = n / sum(n) * 100)
```

## Chronic condition data

```{r read_cc, echo = FALSE, message = FALSE}
# Read chronic conditions data on each year
# Data specification of MSBF files
fts_cc <- read_excel("./Data/mbsf_cc_format.xlsx")

# Create file names
fname <- paste0("./Data/12172/", year, "/mbsf_cc_summary_res000058038_req012172_", year, ".dat")

# Read all MSBF files of 13 years
all_cc <- fname %>% 
  lapply(\(x) read_fwf(x, fwf_widths(fts_cc$length, fts_cc$long_name))) %>% 
  lapply(\(x) anti_join(x, exclude_BENE_IDs)) %>% 
  setNames(year)

# Long format over years
all_cc_long <- all_cc %>% 
  do.call(rbind, .) %>% 
  arrange(BENE_ID, BENE_ENROLLMT_REF_YR)
```

* All chronic condition files of 2008-2020 were imported into R. After excluding duplicated cases described above, we have the same number of observations as in the MSBF file (n = 44,585 unique BeneIDs).
* CC files were merged with MSBF to identify Alzheimer's disease and related dementia cases based on ``ALZH_DEMEN_EVER``: the date when the beneficiary first met the criteria for the disease (after 1999)
  * Among 44,585 subjects, there were 8074 cases (18.1%) of Alzheimer/dementia cases

```{r alz_cases, echo = FALSE}
# Generate data for demographic table
RTI_labels <- c("Unknown", "NH-White", "Black", "Other", "Asian", "Hispanic", "Native")
stop_date  <- as.Date("2008-12-31")

msbf_demog <- all_msbf_long %>% 
  group_by(BENE_ID) %>% 
  slice(n()) %>% 
  select(BENE_ID, AGE_AT_END_REF_YR:ESRD_IND) %>% 
  mutate(BENE_BIRTH_DT = ymd(BENE_BIRTH_DT),
         BENE_DEATH_DT = ymd(BENE_DEATH_DT),
         Dead = ifelse(is.na(BENE_DEATH_DT), 0, 1),
         Dead = factor(Dead, levels = 0:1, labels = c("Alive", "Dead")),
         COVSTART      = ymd(COVSTART),
         SEX_IDENT_CD  = factor(SEX_IDENT_CD, levels = 1:2, labels = c("Male", "Female")),
         RTI_RACE_CD   = factor(RTI_RACE_CD, levels = 0:6, labels = RTI_labels),
         Age_2008      = trunc((BENE_BIRTH_DT %--% stop_date) / years(1)),
         AgeCat_2008 = cut(Age_2008, breaks = c(0, 3:11 * 10), right = FALSE))

cc_alzh <- all_cc_long %>% 
  mutate(ALZH_DEMEN_YN = ifelse(is.na(ALZH_DEMEN_EVER) & is.na(ALZH_EVER), 0, 1),
         ALZH_DEMEN_YN = factor(ALZH_DEMEN_YN, label = c("No", "Yes"))) %>% 
  group_by(BENE_ID) %>% 
  slice(n()) %>%
  select(BENE_ID, ALZH_DEMEN_YN, ALZH_EVER, ALZH_DEMEN_EVER)

cc_alzh %>% 
  group_by(ALZH_DEMEN_YN) %>% 
  tally() %>% 
  mutate(pct = n / sum(n) * 100)
```

* Number of diagnosis by year
  * Notice that some of these cases are prevalent cases before the AHS-2 baseline questionnaire

```{r alz_years, echo = FALSE}
all_cc_long %>% 
  filter(!is.na(ALZH_DEMEN_EVER)) %>% 
  mutate(ALZH_DEMEN_DX_YR = substr(ALZH_DEMEN_EVER, 1, 4)) %>% 
  group_by(BENE_ID) %>% 
  slice(n()) %>% 
  group_by(ALZH_DEMEN_DX_YR) %>% 
  tally() %>% 
  mutate(pct = n / sum(n) * 100) %>% 
  print(n = Inf)
```

## Descriptive tables

### Overall
```{r descriptive_all, echo = FALSE}
msbf_demog %>% 
  left_join(cc_alzh, by = "BENE_ID") %>% 
  CreateTableOne(vars = c("SEX_IDENT_CD", "Age_2008", "RTI_RACE_CD", "Dead"), data = .) %>% 
  print(showAllLevels = TRUE) %>% 
  kableone()
```

### Stratified by Alzheimer/dementia cases

```{r descriptive_stratified, echo = FALSE}
msbf_demog %>% 
  left_join(cc_alzh, by = "BENE_ID") %>% 
  CreateTableOne(strata = "ALZH_DEMEN_YN", vars = c("SEX_IDENT_CD", "Age_2008", "RTI_RACE_CD", "Dead"), data = .) %>% 
  print(showAllLevels = TRUE) %>% 
  kableone()
```

## Cost and utilization data

```{r read_cu, echo = FALSE, message = FALSE}
# Read Cost Use data on each year
# Data specification of CU files
fts_cu <- read_excel("./Data/mbsf_cu_format.xlsx")

# Create file names
fname <- paste0("./Data/12172/", year, "/mbsf_costuse_res000058038_req012172_", year, ".dat")

# Read all MSBF files of 13 years
all_cu <- fname %>% 
  lapply(\(x) read_fwf(x, fwf_widths(as.integer(fts_cu$length), fts_cu$long_name))) %>% 
  lapply(\(x) anti_join(x, exclude_BENE_IDs)) %>% 
  setNames(year)

# Long format over years
all_cu_long <- all_cu %>% 
  do.call(rbind, .) %>% 
  arrange(BENE_ID, BENE_ENROLLMT_REF_YR)

# Unique beneficiaries across years: n = 44,585
all_cu_bene_ids <- all_cu_long %>% 
  select(BENE_ID) %>% 
  distinct()
```

* All CU files of 2008-2020 were imported into R. After excluding duplicated cases described above, we have the same number of observations as in the MSBF file (n = 44,585 unique BeneIDs).

## MedPAR (Medicare provider analysis and review) data

```{r read_mp, echo = FALSE, message = FALSE, warning = FALSE}
# Read MEDPAR data on each year
# Data specification of MEDPAR files
fts_mp <- read_excel("./Data/mbsf_mp_format.xlsx")

# Create file names
fname <- paste0("./Data/12172/", year, "/medpar_all_file_res000058038_req012172_", year, ".dat")

# Read all MSBF files of 13 years
all_mp <- fname %>% 
  lapply(\(x) read_fwf(x, fwf_widths(fts_mp$length, fts_mp$long_name))) %>% 
  lapply(\(x) anti_join(x, exclude_BENE_IDs)) %>% 
  setNames(year)

# Long format over years
all_mp_long <- all_mp %>% 
  do.call(rbind, .) %>% 
  arrange(BENE_ID, MEDPAR_YR_NUM)

# Unique beneficiaries across years: n = 25,125
all_mp_bene_ids <- all_mp_long %>% 
  select(BENE_ID) %>% 
  distinct()
```

* All MedPAR files of 2008-2020 were imported into R. After excluding duplicated cases described above, there are n = 25,125 unique BeneIDs during 13 years.
  * All of these BeneIDs do appear in the MSBF file.
